### 2017-03-24 project_plan/apt-project.tex
Project plan added

\documentclass[times, utf8]{article}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{listings}
\usepackage{xcolor}
\newcommand\myworries[1]{\textcolor{red}{#1}}
\usepackage{hyperref} 
\usepackage{natbib}
\usepackage{subfig}
\usepackage{color}
\usepackage{placeins}
\usepackage[T1]{fontenc}

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}

\lstset{ %
  backgroundcolor=\color{white},   % choose the background color
  basicstyle=\footnotesize,        % size of fonts used for the code
  breaklines=true,                 % automatic line breaking only at whitespace
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{mygreen},    % comment style
  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  keywordstyle=\color{blue},       % keyword style
  stringstyle=\color{mymauve},     % string literal style
}

\begin{document}

\title{\textbf{Semantic textual similarity}}
\date{March, 2017}
\author{Bruno Gavranovi\'c, Neven Miculini\'c and Stipan Mikuli\'c}
\maketitle

\section*{Topic description}

Semantic textual similarity (STS) measures how similar two text snippets are. The task is, given two sentences, to determine their similarity score on a continuous scale from 0 to 5. The score of 5 denotes a perfect semantic equivalence. This task was offered in multiple tracks, covering various cross-lingual and monolingual pairs. Your task, however, is the track concerned with English monolingual pairs. Of course, you are encouraged to give a shot on other tracks as well.
\section*{Project plan}
First and foremost we want to merge together the data from different sources and build vocabulary. We plan to develop simple baseline first and then build upon that.
After reading referenced papers we decided that our pipeline will look something like the following:

\begin{enumerate}
\item Data preprocessing

\begin{itemize}
\item Tokenization
\item POS tagging
\item Parsing (syntax)
\item ...
\end{itemize}

\item Transforming data to corresponding input for models
\begin{itemize}
\item N-gram overlap
\item TF-IDF vector
\item Paragraph2Vec
\end{itemize}

\item Building different models
\begin{itemize}
\item LSTM neural network
\item SVM
\item Linear Regression
\end{itemize}

\item Evaluation
\end{enumerate}

We will make baseline model as centroid vector of word embeddings from sentence.
\end{document}